### 2014-04-30 doc/thesis/cases.tex
Starting to type Lua standard libraries

\section{Lua standard libraries}

\subsection{Basic functions}

\begin{enumerate}
\item assert (v [, message]) :
$\Top \times
(\String \cup \Nil) \times
\TopStar \rightarrow
\NilStar \cup
(\Top \times
(\String \cup \Nil) \times
\TopStar)$ 
\\
When \textbf{v} is $\Nil$ or $\False$, \textbf{assert}
shows the error \textbf{message}.
Otherwise, \textbf{assert} returns \textbf{v} and all the extra
parameters.
\item collectgarbage ([opt [, arg]]) :
$(\String \cup \Nil) \times
(\Number \cup \Nil) \times
\TopStar \rightarrow
(\Number \times \Nil \times \NilStar) \cup
(\Number \times \Number \times \NilStar) \cup
(\Boolean \times \Nil \times \NilStar)$
\\
The return type depends on the type of \textbf{opt}:
\begin{itemize}
\item $\Nil \times \TopStar \rightarrow \Number \times \NilStar$
\item $``collect" \times \TopStar \rightarrow \Number \times \NilStar$
\item $``stop" \times \TopStar \rightarrow \Number \times \NilStar$
\item $``restart" \times \TopStar \rightarrow \Number \times \NilStar$
\item $``count" \times \TopStar \rightarrow \Number \times \Number \times \NilStar$
\item $``step" \times \TopStar \rightarrow \Boolean \times \NilStar$
\item $``setpause" \times \TopStar \rightarrow \Number \times \NilStar$
\item $``setstepmul" \times \TopStar \rightarrow \Number \times \NilStar$
\item $``isrunning" \times \TopStar \rightarrow \Boolean \times \NilStar$
\item $``generational" \times \TopStar \rightarrow \Number \times \NilStar$
\item $``incremental" \times \TopStar \rightarrow \Number \times \NilStar$
\end{itemize}
\item dofile ([filename]) :
$(\String \cup \Nil) \times
\TopStar \rightarrow
\AnyStar
$
\item error (message [, level]) :
$\String \times
(0 \cup 1 \cup 2 \cup \Nil) \times
\TopStar \rightarrow
\NilStar$
\item \_G : $\{ \String:\Top \}$
\item getmetatable (object) :
$\mu X.\{ \_\_metatable:\Top \} \times
\TopStar \rightarrow
(\Nil \cup \{ \_\_metatable:\Top \}) \times
\NilStar$
\item ipairs (t) :
$\{ \Number:\Any \} \times
\TopStar \rightarrow
\Number \times
\Any \times
\NilStar$
\begin{itemize}
\item What happens when \textbf{t} has a metamethod \textbf{\_\_ipairs}?
\item Will $\{ ``tag":\String,\; \Number:\Any \}$ type check?
\item In fact, \textbf{ipairs} and \textbf{pairs} return three values.
Need to check this.
\end{itemize}
\item load (ld [, source [, mode [, env]]]) :
$(\String \cup (\TopStar \rightarrow \TopStar)) \times
(\String \cup \Nil) \times
(``b" \cup ``t" \cup ``bt" \cup \Nil) \times
(\Top \cup \Nil) \times
\TopStar \rightarrow
((\TopStar \rightarrow \TopStar) \times \NilStar) \cup
(\Nil \times \String \times \NilStar)$
\item loadfile ([filename [, mode [, env]]]) :
$(\String \cup \Nil) \times
(``b" \cup ``t" \cup ``bt" \cup \Nil) \times
(\Top \cup \Nil) \times
\TopStar \rightarrow
((\TopStar \rightarrow \TopStar) \times \NilStar) \cup
(\Nil \times \String \times \NilStar)$
\item next (table [, index]) :
$\{ \Any:\Any \} \times
(\Any \cup \Nil) \times
\TopStar \rightarrow
(\Nil \times \NilStar) \cup
(\Any \times \Any \times \NilStar)$
\item pairs (t) :
$\{ \Any:\Any \} \times
\TopStar \rightarrow
\Any \times 
\Any \times
\NilStar$
\item pcall (f [, arg1, ···]) :
$(\TopStar \rightarrow \TopStar) \times
\TopStar \rightarrow
(\Boolean \times \AnyStar) \cup
(\Boolean \times \String \times \NilStar)$
\item print (···) : $\TopStar \rightarrow \NilStar$
\item rawequal (v1, v2) :
$\Top \times
\Top \times
\TopStar \rightarrow
\Boolean \times
\NilStar$
\item rawget (table, index) :
$\{ \Top:\Top \} \times
\Top \times
\TopStar \rightarrow
\Any \times
\NilStar$
\item rawlen (v) :
$(\{ \Top:\Top \} \cup \String) \times
\TopStar \rightarrow
\Number \times
\NilStar$
\item rawset (table, index, value) :
$\{ \Top:\Top \} \times
\Top \times
\Top \times
\TopStar \rightarrow
\{ \Any:\Any \} \times
\NilStar$
\\
In fact, \textbf{index} cannot be $\Nil$ or \texttt{NaN}.
\item select (index, ···) :
$\Number \times
\AnyStar \rightarrow
\AnyStar$
\\
When \textbf{index} is the string ``\texttt{\#}", the typing is:
$\String \times \AnyStar \rightarrow \Number \times \NilStar$
\item setmetatable (table, metatable) :
$\{ \Top:\Top \} \times
(\Nil \cup \{ \_\_index:\mu X.\{ \Top:\Top \} \}) \times
\TopStar \rightarrow
\{ \_\_index:\mu X.\{ \Top:\Top \} \} \times
\NilStar$
\\
Even though \textbf{metatable} can be $\Nil$, it is not optional.
\item tonumber (e [, base]) :
$\String \times
(\Number \cup \Nil) \times
\TopStar \rightarrow
\Number \times
\NilStar$
\item tostring (v) :
$\Top \times
\TopStar \rightarrow
\String \times
\NilStar$
\item type (v) :
$\Top \times
\TopStar \rightarrow
(``nil" \cup
``number" \cup
``string" \cup
``boolean" \cup
``table" \cup
``function" \cup
``thread" \cup
``userdata") \times
\NilStar$
\item \_VERSION : $\String$
\item xpcall (f, msgh [, arg1, ···]) :
$(\TopStar \rightarrow \TopStar) \times
\Top \times
\TopStar \rightarrow
(\Boolean \times \AnyStar) \cup
(\Boolean \times \String \times \NilStar)$
\end{enumerate}

\subsection{Coroutine manipulation}

\subsection{Modules}

\subsection{String manipulation}

\begin{enumerate}
\item string.byte (s [, i [, j]]) :
$\String \times
(\Number \cup \Nil) \times
(\Number \cup \Nil) \times
\TopStar \rightarrow
\Number \times
\Number{*}$
\item string.char (···) :
$\Number{*} \rightarrow
\String \times
\NilStar$
\item string.dump (function) :
$(\TopStar \rightarrow \TopStar) \times
\TopStar \rightarrow
\String \times
\NilStar$
\item string.find (s, pattern [, init [, plain]]) :
$\String \times
\String \times
(\Number \cup \Nil) \times
(\Boolean \cup \Nil) \times
\TopStar \rightarrow
(\Number \times \Number \times \NilStar) \cup
(\Nil \times \NilStar)$
\item string.format (formatstring, ···) :
$\String \times
\TopStar \rightarrow
\String \times
\NilStar$
\item string.gmatch (s, pattern) :
$\String \times
\String \times
\TopStar \rightarrow
\String{*}$
\\
This function actually returns an iterator,
so I still need to confirm its return type.
\item string.gsub (s, pattern, repl [, n]) :
$\String \times
\String \times
(\String \cup \{ \Top:\Top \} \cup (\TopStar \rightarrow \TopStar)) \times
(\Number \cup \Nil) \times
\TopStar \rightarrow
\String \times
\Number \times
\NilStar$
\item string.len (s) :
$\String \times
\TopStar \rightarrow
\Number \times
\NilStar$
\item string.lower (s) :
$\String \times
\TopStar \rightarrow
\String \times
\NilStar$
\item string.match (s, pattern [, init]) :
$\String \times
\String \times
(\Number \cup \Nil) \times
\TopStar \rightarrow
(\String \times \String{*}) \cup
(\Nil \times \NilStar)$
\item string.rep (s, n [, sep]) :
$\String \times
\Number \times
(\String \cup \Nil) \times
\TopStar \rightarrow
\String \times
\NilStar$
\item string.reverse (s) :
$\String \times
\TopStar \rightarrow
\String \times
\NilStar$
\item string.sub (s, i [, j]) :
$\String \times
\Number \times
(\Number \cup \Nil) \times
\TopStar \rightarrow
\String \times
\NilStar$
\item string.upper (s) :
$\String \times
\TopStar \rightarrow
\String \times
\NilStar$
\end{enumerate}

\subsection{Table manipulation}

\begin{enumerate}
\item table.concat (list [, sep [, i [, j]]]) :
$\{ \Number:\String \cup \Number \} \times
(\String \cup \Nil) \times
(\Number \cup \Nil) \times
(\Number \cup \Nil) \times
\TopStar \rightarrow
\String \times
\NilStar$
\item table.insert (list, [pos,] value) :
$\{ \Number:\Top \} \times
(\Number \times \Top) \cup
(\Top \times \Nil) \times
\TopStar \rightarrow
\NilStar$
\\
This function does not accept/discard extra parameters.
Perhaps due to the fact that the parameter \textbf{pos} is optional.
If \textbf{value} is $\TopStar$, the function inserts \textbf{pos}
in the \textbf{list}, that is, the input type is
$\{ \Number:\Top \} \times \Top \times \TopStar$.
If \textbf{value} is $\Nil$, the function does not insert \textbf{pos}
in the \textbf{list}, that is, the input type is
$\{ \Number:\Top \} \times \Number \times \Top \times \TopStar$,
but the element is not inserted in the list because $\Top$ is actually $\Nil$.
\item table.pack (···) :
$\TopStar \rightarrow
\{ \Number:\Top \} \times
\NilStar$
\\
Even though \textbf{table.pack} gets any number of parameters of
any type, the $\Nil$ type generates a hole in the returned table.
\item table.remove (list [, pos]) :
$\{ \Number:\Top \} \times
(\Number \cup \Nil) \times
\TopStar \rightarrow
\Top \times
\NilStar
$
\\
The return type depends on the type of the element that is removed
from the list.
\item table.sort (list [, comp]) :
$\{ \Number:\Top \} \times
((\TopStar \rightarrow \TopStar) \cup \Nil) \times
\TopStar \rightarrow
\NilStar$
\item table.unpack (list [, i [, j]]) :
$\{ \Number:\Top \} \times
(\Number \cup \Nil) \times
(\Number \cup \Nil) \times
\TopStar \rightarrow
\TopStar$
\\
The return type depends on the type of the elements that are stored in
the list.
\end{enumerate}

\subsection{Mathematical functions}

\begin{enumerate}
\item math.abs (x) :
$\Number \times \TopStar \rightarrow \Number \times \NilStar$
\item math.acos (x) :
$\Number \times \TopStar \rightarrow \Number \times \NilStar$
\item math.asin (x) :
$\Number \times \TopStar \rightarrow \Number \times \NilStar$
\item math.atan (x) :
$\Number \times \TopStar \rightarrow \Number \times \NilStar$
\item math.atan2 (y, x) :
$\Number \times \Number \times \TopStar \rightarrow \Number \times \NilStar$
\item math.ceil (x) :
$\Number \times \TopStar \rightarrow \Number \times \NilStar$
\item math.cos (x) :
$\Number \times \TopStar \rightarrow \Number \times \NilStar$
\item math.cosh (x) :
$\Number \times \TopStar \rightarrow \Number \times \NilStar$  
\item math.deg (x) :
$\Number \times \TopStar \rightarrow \Number \times \NilStar$
\item math.exp (x) :
$\Number \times \TopStar \rightarrow \Number \times \NilStar$
\item math.floor (x) :
$\Number \times \TopStar \rightarrow \Number \times \NilStar$
\item math.fmod (x, y) :
$\Number \times \Number \times \TopStar \rightarrow \Number \times \NilStar$
\item math.frexp (x) :
$\Number \times \TopStar \rightarrow \Number \times \Number \times \NilStar$
\item math.huge : $\Number$
\item math.ldexp (m, e) :
$\Number \times \Number \times \TopStar \rightarrow \Number \times \NilStar$
\item math.log (x [, base]) :
$\Number \times
(\Number \cup \Nil) \times
\TopStar \rightarrow
\Number \times
\NilStar$
\item math.max (x, ···) :
$\Number \times \Number{*} \rightarrow \Number \times \NilStar$
\item math.min (x, ···) :
$\Number \times \Number{*} \rightarrow \Number \times \NilStar$
\item math.modf (x) :
$\Number \times \TopStar \rightarrow \Number \times \Number \times \NilStar$
\item math.pi : $\Number$
\item math.pow (x, y) :
$\Number \times \Number \times \TopStar \rightarrow \Number \times \NilStar$
\item math.rad (x) :
$\Number \times \TopStar \rightarrow \Number \times \NilStar$
\item math.random ([m [, n]]) :
$(\Number \cup \Nil) \times
(\Number \cup \Nil) \rightarrow
\Number \times
\NilStar$
\\
This function does not accept $\Nil$ as its only parameter,
although the parameter passing is optional.
This function does not accept/discard extra parameters.
\item math.randomseed (x) :
$\Number \times \TopStar \rightarrow \NilStar$
\item math.sin (x) :
$\Number \times \TopStar \rightarrow \Number \times \NilStar$
\item math.sinh (x) :
$\Number \times \TopStar \rightarrow \Number \times \NilStar$
\item math.sqrt (x) :
$\Number \times \TopStar \rightarrow \Number \times \NilStar$
\item math.tan (x) :
$\Number \times \TopStar \rightarrow \Number \times \NilStar$
\item math.tanh (x) :
$\Number \times \TopStar \rightarrow \Number \times \NilStar$
\end{enumerate}

\subsection{Bitwise operations}

\subsection{Input and Output facilities}

\subsection{Operating System facilities}

\subsection{The debug library}