### 2015-03-22 doc/thesis/conclusions.tex
Improving conclusions

In this work we presented Typed Lua, an optional type system for Lua.
Typed Lua is a Lua extension that allows programmers to combine
static and dynamic typing in Lua code, making easier the evolution
of simple scripts into large programs.

Our main contribution is the formalization of a complete optional type
system that introduces several novel type system features to statically
type check some Lua idioms and features.
Even though Lua shares several features with other dynamically
typed languages such as JavaScript, Lua also has several unusual features.
These unusual features include tables (or associative arrays) as the solely data
structure mechanism, functions with multiple return values, and functions
with flexible arity that interact with multiple assignment.
We highlight the following novel features of our type system:
\begin{itemize}
\item the refinement of table types support the incremental evolution
of record and object types, playing an important role to statically
type check the idiomatic way that Lua programmers use tables to
define modules and objects;
\item projection types handle functions that are overloaded on the
number of return values, allowing programmers to narrow the type of
a local variable that depends on the type of another one;
\item union types and variadic types help our type system handling
functions with flexible arity, that is, union types are helpful to
describe optional parameters while variadic types are helpful to
describe the type of the vararg expression and the type of functions
that can receive or return any number of values.
\end{itemize}

A key feature in optional type systems is usability.
This means that optional type systems should not change the idioms
that programmers are already familiar with.
Instead, optional type system should fit existing idioms to
statically type check them.
In other words, design a simple type system that only relies on
the semantics and run-time tags of a dynamically typed language
can overload programmers, forcing them to code in a different way.
Thus, the most challenging aspect of optional type systems is
to design a complex type system that feels natural to the programmers.

Before starting the design of Typed Lua, we took usability into consideration.
We realized that we should not rely on the semantics of Lua only,
as this could lead to a cumbersome type system that would not support
several idiomatic Lua code.
For this reason, we performed a mostly automated survey of Lua idioms
and features to inform our design choices.

After designing and implementing Typed Lua, we performed several
case studies to evaluate the usability of our type system.
Our evaluation results showed that our type system can statically
type check several Lua idioms and features, though the evaluation
results also exposed several limitations of our type system.
We found that the three main limitations of our type system are
the lack of intersection types, parametric polymorphism, and operator overloading.
Overcoming these limitations is our major target for future work,
as it will allow us to statically type check more programs.

Unlike other optional type systems, we designed Typed Lua without
deliberated unsound parts.
However, we still do not have proofs that the novel features of
our type system are sound.
We see a soundness proof as another major future work, as it is
necessary to use static types for code optimization.

Finally, we believe that Typed Lua is a major contribution to the Lua community,
because it offers a framework that programmers can use to document,
test, and better structure their applications.
For libraries where a full conversion to static type checking should
prove unfeasible or too much work, the community can use Typed Lua
just to document the external interfaces of the libraries,
giving the benefits of static type checking to the users of these
libraries.
In fact, we already have user feedback from Lua programmers that are
using Typed Lua in their projects.
For instance, ZeroBrane Studio is an IDE for Lua development that is
starting to use Typed Lua to perform static analysis in Lua code.