### 2015-05-11 doc/thesis/glossary.tex
Defining the glossary terms to be defined

\begin{description}
\item[bottom type] TO DO.
\item[closed table type] TO DO.
\item[coercion] TO DO.
\item[consistency] TO DO.
\item[consistent-subtyping] TO DO.
\item[contravariant] TO DO.
\item[covariant] TO DO.
\item[depth subtyping] TO DO.
\item[dynamic type] TO DO.
\item[filter type] TO DO.
\item[fixed table type] TO DO.
\item[flow typing] TO DO.
\item[free assigned variable] TO DO.
\item[gradual type system] TO DO.
\item[gradual typing] TO DO.
\item[invariant] TO DO.
\item[metatable] TO DO.
\item[nominal type system] TO DO.
\item[open table type] TO DO.
\item[optional type system] TO DO.
\item[projection environment] TO DO.
\item[projection type] TO DO.
\item[prototype object] TO DO.
\item[self-like delegation] TO DO.
\item[sound type system] TO DO.
\item[structural type system] TO DO.
\item[table refinement] TO DO.
\item[top type] TO DO.
\item[type environment] TO DO.
\item[type tag] TO DO.
\item[unique table type] TO DO.
\item[unsound type system] TO DO.
\item[userdata] TO DO.
\item[vararg expression] TO DO.
\item[variadic function] TO DO.
\item[variance] TO DO.
\item[width subtyping] TO DO.
\end{description}