### 2015-08-02 doc/thesis/ack.tex
Adding acks and short resume

First of all I would like to thank three indispensable people in my life:
my wife, my mother, and my father.
Izabella is my inexhaustible source of inspiration, and she is
always pushing me forward.
My mother is a very brave person, and she inspires me to never give up.
My father is a very kind person, and he inspires me to try being more sensitive.
I am thankful to them not only because the support they gave me while
writing this thesis, but also because they are the reason of my life.

This work would have not happened without all the help from my advisors.
Roberto has a singular view of computer science that pushed me to
pursuit the best I could do, while Fabio gave me the confidence I
needed about the work we were doing.
I am glad that both of them had patience to guide me through this
work, and for all the nice meetings we had that led me to learn
a lot of new concepts and ideas.

It is hard to complete a thesis without having some leisure moments,
so I thank my friend and roommate André Ramiro for all the crazy and
fun moments that we had, which helped me to keep focused during working hours.

I also thank my friends at LabLua for the nice work environment,
specially Hisham and Pablo for all the support during the
hardest moments.

It may look easy moving to Rio, as it is a sunny place with lots
of things to do, but it may not be as easy as it looks for
someone that grew up in a cloudy and rainy town.
Thus, I thank my cousin Rebecca and my aunt Nazira for all the help
and affection they gave me during my adaptation in Rio, which
were fundamental to keep my goals in mind.

I thank the guys from Rio de Janeiro Fixed Gear (RJFG) for all
the fun we had during the night rides that I participated,
and for helping me to discover Rio.

I am also most grateful to the professors Ana Lúcia de Moura,
Anamaria Martins Moreira, Edward Hermann Haeusler, and
Roberto da Silva Bigonha for their careful review that helped
to improve this work.

Finally, I would like to thank CAPES, Google Summer of Code, and PUC-Rio
for partially funding this work.